---
title: "FINAL_DRAFT"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


#### Importing Required Packages 
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

package_install_load <- function(x) 
{
  if (!require(x, character.only = TRUE)) 
  {
    install.packages(x, dep = TRUE)
    if (!require(x, character.only = TRUE)) 
    {
      stop('Package not found')
    }
  }
}

## Block - Load Required Packages
packages <- c("rlang","tidyverse","ggplot2","gridExtra","dplyr","sqldf","readxl","readr","tidytext","tidyr"
              ,"maps","lubridate","treemap","stringi","stringr","plyr","leaflet")
invisible(lapply(packages, package_install_load))
```


#### Function - Set Working Directory
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

set_workspace <- function(dir_path) 
{
  
  #dir.create(dir_path) # Create Directory
  setwd(dir_path) #Set Working Directory
  print(paste("Working Directory Set to : ",dir_path)) # Print Message
}

## Setting Workspace
#set_workspace("C:/Users/K/Downloads/DS5110 Project");
set_workspace("C:/Users/aggarwal.k/Downloads/DS5110 Project");

```



### (I) Data Acquisition:
### (a) Acquisition of Data:

```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

loan <- read_csv("kiva_loans.csv")


reg_loc <- read_csv("kiva_mpi_region_locations.csv")


theme <- read_csv("loan_theme_ids.csv")


theme_reg <- read_csv("loan_themes_by_region.csv")

flood_dataset <- read_excel("Flood_Risk.xlsx")


data(world.cities)
world_cities <- world.cities

remove(world.cities)
```

### (b) Glimpse of Data:
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
head(loan)
glimpse(loan)


head(reg_loc)
glimpse(reg_loc)


head(theme)
glimpse(theme)


head(theme_reg)
glimpse(theme_reg)

```




### Summary Statistics Data
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#summary(loan)
#summary(reg_loc)
#summary(theme)
#summary(theme_reg)



#Lets identify the total funded amount by Kiva to the field agents
total_funded_amnt <- sum(loan$funded_amount)
total_funded_amnt

```



### Data Preparation 


### Data Cleaning & Shaping

#### (I) Loan Time Formats
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

#separate funded time into years only, group by years and indentify the analysis
loan <- loan %>% mutate(funding_year = year(funded_time)) 



```



#### (II) Removal and Alteration of Columns
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#Unique observation where kiva has active loans
reg_loc <- unique(reg_loc[, !(colnames(reg_loc) %in% c("geo"))])
loan <- unique(loan[, !(colnames(loan) %in% c("date"))])
theme <- unique(theme[, !(colnames(theme) %in% c("id"))])
#theme_reg_raw <- unique(theme_reg[, !(colnames(theme_reg) %in% c("Partner ID"))])

```


#### (iii) Modifying "borrower_genders" variable to replace every instance with single gender i.e male or female
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
loan <- loan %>% mutate(gender = ifelse(str_detect(borrower_genders, "female"), "female", "male"))

```



### Data Exploration:

### Exploratory Data Plots: General


#### (i) Loan
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#levels of sector
table(loan$sector)

#levels of repayment interval
table(loan$repayment_interval)

#levels of gender
table(loan$gender)

#summary statsistics of funded amount by kiva 
summary(loan$funded_amount)

```



#### (ii) MPI_REGION_LOC
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#Measurement poverty index
summary(reg_loc$MPI)

#identifying which columns with observations across the table is not having missing values
colSums(!is.na(reg_loc))

#identify whuch columns with observations across the table has missing values 
colSums(is.na(reg_loc))

#Lets remove the records where MPI is missing, since there will be no meaning of the other variables
reg_loc <- reg_loc %>%
  filter(!is.na(MPI))

theme_reg$country <- ifelse(theme_reg$country == 'Viet Nam','Vietnam',theme_reg$country)
reg_loc$country <- ifelse(reg_loc$country == 'Viet Nam','Vietnam',reg_loc$country)

theme_reg$country <- ifelse(theme_reg$country == 'Myanmar (Burma)','Myanmar',theme_reg$country)
reg_loc$country <- ifelse(reg_loc$country == 'Myanmar (Burma)','Myanmar',reg_loc$country)

#level of world regions
table(reg_loc$world_region)
```



### Join of loan and reg_loc 
####Benefit, we can map active loans data with world region and measurement poverty index. Alsom will be able to map latitude and longitude of the location.  
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
# colnames(world_cities) <- gsub("\\.", "_", colnames(world_cities))
# reg_loc <- sqldf("select  country_etc as country
#                   ,wc.pop as pop
#                   ,wc.lat,wc.long,world_region,MPI,region from world_cities wc
#                  left join reg_loc rl on wc.country_etc = rl.country")



#natural join by country and region
loan_reg_loc <- loan %>%
  left_join(reg_loc,by = c("country"="country"))


#further perform analysis based on loan_mpi. Show EDA based on this


```




#### (iii) THEME_REG
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#identify which columns with observations across the table has missing values 
colSums(is.na(theme_reg))


#level of sector of field agents
table(theme_reg$sector)


#table(theme_reg$`Loan Theme Type`)

```



#### (iv) THEME
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#identify which columns with observations across the table has missing values 
colSums(is.na(theme))

```


### Join of theme and theme_reg
####Benefit, we can map active loans data with world region and measurement poverty index. Alsom will be able to map latitude and longitude of the location.

```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#natural join by country and region
theme_reg_join <- theme_reg %>%
  left_join(theme,by = c("Partner ID" = "Partner ID"))

#further perform analysis based on loan_mpi. Show EDA based on this
```



### ggplots - General

#### (i) Popularity of sector where large number of customers got funded
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
loan %>%
  group_by(sector) %>%
  dplyr::summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  mutate(sector = reorder(sector,Count)) %>%
  ggplot(aes(x = sector,y = Count)) +
  geom_bar(stat = "identity", fill = "Gold") +
  geom_text(aes(x = sector, y = 2, label = paste0("(",Count,")",sep=""))
            , hjust = -.1, vjust = .3, fontface = "bold") +
  xlab("Sector") +
  ylab("Frequency") +
  ggtitle("Popularity of Sectors in Funding") +
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5))

```




#### (ii) Popularity of sector where large number of funds been given
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
#Box-Plot, since using categorical against continuous

loan %>%
      mutate(fill = as.factor(sector))%>%
      ggplot(aes(x = sector, y= log2(funded_amount), fill = sector)) +
      geom_boxplot() +
      labs(x= 'Sector',y = 'Funded Amount', 
           title = paste0("Distribution of", ' Funded Amount ')) +
       theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(plot.title = element_text(hjust = 0.5))
```



#### (iii) Identifying the most number of repayment interval
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

loan %>%
  group_by(repayment_interval) %>%
  dplyr::summarise(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(repayment_interval = reorder(repayment_interval,count)) %>%
  ggplot(aes(x = repayment_interval,y = count)) +
  geom_bar(position = position_dodge(), stat = "identity", fill = "green") +
  geom_text(aes(x = repayment_interval, y = 2, label = paste0("(",count,")",sep="")), hjust = -.1, vjust =   .3, fontface = "bold") +
  xlab("Repayment Interval") +
  ylab("Frequency") +
  ggtitle("Repayment Interval Frequency") +
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5))

```



#### (iv) Identifying the gender to whom the loans been given
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

loan %>%
  filter(!is.na(gender)) %>%
  group_by(gender) %>%
  dplyr::summarise(count = n()) %>%
  arrange(desc(count)) %>%
  mutate(gender = reorder(gender,count)) %>%
  ggplot(aes(x = gender,y = count/1000)) +
  geom_bar(position = position_dodge(), stat = "identity", fill = "skyblue") +
  geom_text(aes(x = gender, y = 2, label = paste0("(",count/1000,")",sep="")), hjust = -.1, vjust = .3, fontface = "bold") +
  xlab("Gender") +
  ylab("No. of Loans (x1000)") +
  ggtitle("Loan Distribution by Gender") +
  coord_flip()+
  theme(plot.title = element_text(hjust = 0.5))

```


#### (v) Popularity of the activity where kiva has active loans
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}

loan %>%
  group_by(activity) %>%
  dplyr::summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  ungroup() %>%
  mutate(activity = reorder(activity,Count)) %>%
  top_n(10) %>%
  
  ggplot(aes(x = activity,y = Count)) +
  geom_bar(stat='identity',colour="white", fill = "light coral") +
  geom_text(aes(x = activity, y = 1, label = paste0("(",Count,")",sep="")),
            hjust=0, vjust=.5, size = 4, colour = 'black',
            fontface = 'bold') +
  labs(x = 'Activity', 
       y = 'Frequency', 
       title = 'Activities for Which Loan was Procured') +
  coord_flip() +
   theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

```

#### (vi) Popularity by Country where KIVA has active loans
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
loan %>%
  group_by(country) %>%
  dplyr::summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  ungroup() %>%
  mutate(country = reorder(country,Count)) %>%
  head(10) %>%
  
  ggplot(aes(x = country,y = Count)) +
  geom_bar(stat='identity',colour="white", fill = "light green") +
  geom_text(aes(x = country, y = 1, label = paste0("(",Count,")",sep="")),
            hjust=0, vjust=.5, size = 4, colour = 'black',
            fontface = 'bold') +
  labs(x = 'Country', 
       y = 'Frequency', 
       title = 'Active Loans by Country') +
  coord_flip() +
   theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))

```
#### The following plot shows the most popular themes in a bar chart. We have removed rows where the theme was not mentioned.

#### General is the most popular theme which does not give us a lot of information.

#### Underserved is the next popular theme, followed by Agriculture, Rural Inclusion, Water and Higher Education


#### (vii) Popularity of Loan Themes
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
theme %>%
  dplyr::rename (themeType = `Loan Theme Type`) %>%
  filter(!is.na(themeType)) %>%
  group_by(themeType) %>%
  dplyr::summarise(Count = n()) %>%
  arrange(desc(Count)) %>%
  ungroup() %>%
  mutate(themeType = reorder(themeType,Count)) %>%
  head(10) %>%
  
  ggplot(aes(x = themeType,y = Count)) +
  geom_bar(stat='identity',colour="white", fill = "purple") +
  geom_text(aes(x = themeType, y = 1, label = paste0("(",Count,")",sep="")),
            hjust=0, vjust=.5, size = 4, colour = 'black',
            fontface = 'bold') +
  labs(x = 'Type of Theme', 
       y = 'Frequency', 
       title = 'Themes for Which Loan was Procured') +
  coord_flip() +
   theme_bw()+
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
colnames(world_cities) <- gsub("\\.", "_", colnames(world_cities))

world_cities_ctry <- sqldf("select country_etc as country,median(lat) as lat, median(long) as lon
                           from world_cities
                           group by country_etc")


country_loans <- sqldf("select wc.country,sum(funded_amount) amount,wc.lat,wc.lon,avg(MPI) MPI
                      from loan cl
                      inner join world_cities_ctry wc on wc.country = cl.country
                      left join reg_loc rl on rl.country = wc.country
                       group by wc.country,wc.lat,wc.lon")
```

## The below function generates a map of the region assigned and plots the loans disbursed 

```{r}
plotmap_by_country <- function(country_loans)
{
  
  center_lon = median(country_loans$lon,na.rm = TRUE)
  center_lat = median(country_loans$lat,na.rm = TRUE)
  
  
  leaflet(country_loans) %>% addTiles() %>%
    addCircles(lng = ~lon, lat = ~lat,radius = ~(amount/1000) ,
               color ="Blue")  %>%
    # controls
    setView(lng=center_lon, lat=center_lat,zoom = 2) 
}
```


## The function below creates plots for indicating countries top 10 countries in a region where maximum loan has been disbursed

```{r}
  country_loans %>%
    group_by(country) %>%
    dplyr::summarise(tot_amt = sum(amount)) %>%
    arrange(desc(tot_amt)) %>%
    ungroup() %>%
    mutate(country = reorder(country,tot_amt)) %>%
    head(10) %>%
    
    ggplot(aes(x = country,y = tot_amt/1000)) +
    geom_bar(stat='identity',colour="white", fill = "Light coral") +
    geom_text(aes(x = country, y = 1, label = paste0("(",tot_amt/1000,")",sep="")),
              hjust=0, vjust=.5, size = 4, colour = 'black',
              fontface = 'bold') +
    labs(x = 'Countries', 
         y = 'Loan Amount (x1000)', 
         title = 'Loan Amount by Countries') +
    coord_flip() +
    theme_bw()+
    theme(plot.title = element_text(hjust = 0.5))
```

## The Map below shows the areas of Africa were KIVA loans have been disbursed
```{r}
plotmap_by_country(country_loans)
```

## Tree Map of the funded loan amount by Country 
```{r, message=FALSE, warning=FALSE, paged.print=FALSE}
loans_funded_amount = country_loans %>%
  group_by(country) %>%
  dplyr::summarise(tot_amt = sum(amount)) %>%
  arrange(desc(tot_amt)) %>%
  ungroup() %>%
  mutate(country = reorder(country,tot_amt)) %>%
  head(20) 


treemap(loans_funded_amount, 
        index="country", 
        vSize = "tot_amt",  
        title="Funded Amount", 
        palette = "RdBu",
        fontsize.title = 14 
)
```

## Flood Calamities Analysis

```{r}
flood_dataset <- read_excel("Flood_Risk.xlsx")


flood_risks <- flood_dataset %>%
  left_join(country_loans, by=c("Country"="country")) %>%
  mutate(tot_amt = sum(country_loans$amount)) %>%
  mutate(Percentage_Prop = (amount/tot_amt)*100) %>%
  select("Country","Affected_Pop","amount","MPI","Percentage_Prop")


top_10_flood_risks_funding <- flood_risks %>%
  arrange(desc(Affected_Pop)) %>%
  head(10) 

```


```{r}
top_10_flood_risks_funding$Percentage_Prop <- ifelse(top_10_flood_risks_funding$Percentage_Prop > 0 ,top_10_flood_risks_funding$Percentage_Prop,0)
top_10_flood_risks_funding$Percentage_Prop[is.na(top_10_flood_risks_funding$Percentage_Prop)] <- 0



scaleFactor <- max(top_10_flood_risks_funding$Affected_Pop/1000) / max(top_10_flood_risks_funding$Percentage_Prop)


ggplot(top_10_flood_risks_funding, aes(x=Country,  width=.4)) +
  geom_col(aes(y=Affected_Pop/1000), fill="blue", position = position_nudge(x = -.4)) +
  geom_col(aes(y=Percentage_Prop * scaleFactor), fill="red") +
  scale_y_continuous(name="Affected Population (x1000)", sec.axis=sec_axis(~./scaleFactor, name="%age of Kiva Loans")) +
  theme(
    axis.title.y.left=element_text(color="blue"),
    axis.text.y.left=element_text(color="blue"),
    axis.title.y.right=element_text(color="red"),
    axis.text.y.right=element_text(color="red")
  ) +
  labs(title = "Kiva Loans vs Flood Affected Population", x = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme(axis.title.y = element_text(vjust = 0.6)) +
  theme(axis.text.x =
                   element_text(size  = 10,
                                angle = 45,
                                hjust = 1,
                                vjust = 1)) 
  
```
```{r}
flood_risks %>%
  ggplot(aes(x=log(Affected_Pop/100000))) +
  geom_point(aes(y=(amount/10000)))


flood_risks %>%
  ggplot(aes(x=log(MPI^2))) +
  geom_point(aes(y=amount/10000))
```